---
title: "CEDEN Benthic Data"
author: "Eric Lawrence"
date: "12/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)
```

## CEDEN Data 

This document covers the data analysis for the benthic macroinvertebrate dataset obtained from the CEDEN Advanced Query Tool: 

    https://ceden.waterboards.ca.gov/AdvancedQueryTool

I downloaded the CEDEN Benthic data with the following parameters:

   Region selection by County: Contra Costa, Sacramento, San Joaquin, Solano, and Yolo.

   Date Range: 1/1/2010 to 6/19/2019 (latest available date).

I also used the CEDEN water quality data from the Tox Box accessed on 12/13/2020:

   Z:\\Upper San Francisco Project\\Data & Analyses\\Year 2\\CEDEN Files by Topic\\CEDEN Water WQ Data.xlsx

## Risk Regions
   
To assign risk regions to the data with spatial coordinates, I used the RiskRegions_DWSC_Update_9292020.shp shapefile obtained from the Tox Box.


## Load in libraries and data

```{r libraries and data}

library(tidyverse)
library(sf)
library(here)
library(readxl)

USFE.riskregions <- here("data/RiskRegions_DWSC_Update_9292020.shp") %>% # path relative to main directory
  st_read()

st_crs(USFE.riskregions)

crs.WGS84 <- st_crs(USFE.riskregions)


#### Load in CEDEN data

CEDEN <- read_excel("data/ceden_data_edited.xlsx")
```

