---
title: "CEDEN Benthic Data"
author: "Eric Lawrence"
date: "12/13/2020"
output:
  html_document:
    code_folding: show
    code_download: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE, warning = FALSE, message = FALSE)
```

Example text yay

## CEDEN Data 

This document covers the data analysis for the benthic macroinvertebrate dataset obtained from the CEDEN Advanced Query Tool: 

    https://ceden.waterboards.ca.gov/AdvancedQueryTool

I downloaded the CEDEN Benthic data with the following parameters:

   Region selection by County: Contra Costa, Sacramento, San Joaquin, Solano, and Yolo.

   Date Range: 1/1/2010 to 6/19/2019 (latest available date).

I also used the CEDEN water quality data from the Tox Box accessed on 12/13/2020:

   Z:\\Upper San Francisco Project\\Data & Analyses\\Year 2\\CEDEN Files by Topic\\CEDEN Water WQ Data.xlsx

## Risk Regions
   
To assign risk regions to the data with spatial coordinates, I used the RiskRegions_DWSC_Update_9292020.shp shapefile obtained from the Tox Box.


## Load in libraries and data

```{r, echo=TRUE, results='hide'}

library(tidyverse)
library(sf)
library(here)
library(readxl)

# Load risk regions shapefile
USFE.riskregions <- here("data/RiskRegions_DWSC_Update_9292020.shp") %>%
  st_read()

# Create WGS84 projection variable
crs.WGS84 <- st_crs(USFE.riskregions)

#Load in CEDEN data
CEDEN <- read_excel("data/ceden_data_edited.xlsx")
```

## Set CEDEN Projection to WGS84

The CEDEN data has data with different coordinate systems depending on the source project. I separated the projects by coordinate systems into separate dataframes, converted the dataframes to WGS84 seperately, then recombined them. There were some data that had no known coordinate system so those were excluded from the analysis.

EPSG numbers for different coordinate systems can be found at https://spatialreference.org/ref/epsg/

```{r echo=TRUE, results='hide'}
### Separate by Datum
unique(CEDEN$Datum)

ceden.wgs84 <- filter(CEDEN, Datum == "WGS84")
ceden.nad83 <- filter(CEDEN, Datum == "NAD83")
ceden.nr <- filter(CEDEN, Datum == "NR")
ceden.NAD27 <- filter(CEDEN, Datum == "NAD27")

### set Datum, shapefiles

ceden.wgs84.sf <- ceden.wgs84 %>%
  st_as_sf(coords=c("TargetLongitude", "TargetLatitude"), crs = crs.WGS84)

ceden.nad83.sf <- ceden.nad83 %>%
  st_as_sf(coords=c("TargetLongitude", "TargetLatitude"), crs = 4269)

ceden.NAD27.sf <- ceden.NAD27 %>%
  st_as_sf(coords=c("TargetLongitude", "TargetLatitude"), crs = 4267)

# ceden.nr does not have a known coordinate system

### Transform into WGS84

ceden.nad83.t.wgs84.sf <- st_transform(ceden.nad83.sf, crs.WGS84)
st_crs(ceden.nad83.t.wgs84.sf)

ceden.NAD27.t.wgs84.sf <- st_transform(ceden.NAD27.sf, crs.WGS84)
st_crs(ceden.NAD27.t.wgs84.sf)


### Join Excel Data to Risk Regions
ceden.wgs84.sf <- st_join(ceden.wgs84.sf, USFE.riskregions["Subregion"])
ceden.nad83.t.wgs84.sf <- st_join(ceden.nad83.t.wgs84.sf, USFE.riskregions["Subregion"])
ceden.NAD27.t.wgs84.sf <- st_join(ceden.NAD27.t.wgs84.sf, USFE.riskregions["Subregion"])

### Combine datasets
ceden.all.sf <- bind_rows(ceden.wgs84.sf, ceden.nad83.t.wgs84.sf)
ceden.all.sf <- bind_rows(ceden.all.sf, ceden.NAD27.t.wgs84.sf)

### Remove records outside of Risk Regions

ceden.all.sf <- ceden.all.sf %>%
  filter(!is.na(Subregion))

```

## Plot Ceden Data and Write csv

```{r, echo=TRUE }

# Basic plot
ggplot() +
  geom_sf(data = USFE.riskregions) +
  geom_sf(data = ceden.all.sf, aes(color = Subregion)) +
  scale_color_brewer(palette = "Set1") + # not color-blind safe
  ggtitle("Ceden Benthic Data")

# Write table

write.csv(ceden.all.sf, "data/ceden_with_RR.csv")


```

# Analysis

## Organize Data

I built a dataframe that grouped the data by station code and sample date so that each record is a sampling event. I summarized the number of taxa and counts by selected Orders and all Phylums. I also calculated the EPT and ETO indicies and taxa counts.


I selected the orders most commonly used in community indicies.

Orders: 

* Ephemeroptera (n_E)
* Plecoptera (n_P)
* Trichoptera (n_T)
* Odonata (n_O)
* Diptera (n_D)


```{r, echo=TRUE, results='hide'}
##### Create dataframe set into groups by station and sample date to isolate individual samples.
##### summarise by taxa presence, number of taxa present, and indexes

samp.df <- ceden.all.sf %>%
  group_by(StationCode, SampleDate) %>%
  summarise(Subregion = first(Subregion),
            StationName = first(StationName),
            Project = first(Project),
            Projectcode = first(Projectcode),
            n_taxa = n(),
            n_E = sum(Orders == "Ephemeroptera", na.rm = TRUE),
            n_P = sum(Orders == "Plecoptera", na.rm = TRUE),
            n_T = sum(Orders == "Trichoptera", na.rm = TRUE),
            n_O = sum(Orders == "Odonata", na.rm = TRUE),
            n_D = sum(Orders == "Diptera", na.rm = TRUE),
            n_Phylum_NA = sum(is.na(Phylum)),
            n_Arthropoda = sum(Phylum == "Arthropoda", na.rm = TRUE),
            n_Annelida = sum(Phylum == "Annelida", na.rm = TRUE),
            n_Nematoda = sum(Phylum == "Nematoda", na.rm = TRUE),
            n_Ectoprocta = sum(Phylum == "Ectoprocta", na.rm = TRUE),
            n_Bacillariophyta = sum(Phylum == "Bacillariophyta", na.rm = TRUE),
            n_Cryptophyta = sum(Phylum == "Cryptophyta", na.rm = TRUE),
            n_Heterokontophyta = sum(Phylum == "Heterokontophyta", na.rm = TRUE),
            n_Ochrophyta = sum(Phylum == "Ochrophyta", na.rm = TRUE),
            n_Coelenterata = sum(Phylum == "Coelenterata", na.rm = TRUE),
            n_Nemertea = sum(Phylum == "Nemertea", na.rm = TRUE),
            n_Mollusca = sum(Phylum == "Mollusca", na.rm = TRUE),
            n_Platyhelminthes = sum(Phylum == "Platyhelminthes", na.rm = TRUE),
            n_Bryozoa = sum(Phylum == "Bryozoa", na.rm = TRUE),
            n_Cyanobacteria = sum(Phylum == "Cyanobacteria", na.rm = TRUE),
            n_Chlorophyta = sum(Phylum == "Chlorophyta", na.rm = TRUE),
            n_Euglenozoa = sum(Phylum == "Euglenozoa", na.rm = TRUE),
            n_Streptophyta = sum(Phylum == "Streptophyta", na.rm = TRUE),
            n_Rhodophyta = sum(Phylum == "Rhodophyta", na.rm = TRUE),
            n_Chordata = sum(Phylum == "Chordata", na.rm = TRUE)
            ) %>%
  mutate(EPT_taxa = sum(n_E, n_P, n_T)) %>%  ### Calculate indexes using mutate
  mutate(EPT_index = EPT_taxa / n_taxa) %>%
  mutate(ETO_taxa = sum(n_E, n_O, n_T)) %>%
  mutate(ETO_index = ETO_taxa / n_taxa)

```

```{r, echo=TRUE}

tibble(samp.df)

```

## Dataframe with one record for station

This dataframe is used for mapping.

```{r, echo=TRUE, results='hide'}

st.df <- ceden.all.sf %>%
  group_by(StationCode) %>%
  summarise(n(), n_distinct(SampleDate), first(Subregion)) %>%
  rename(Subregion = `first(Subregion)`)

```

```{r, echo=TRUE}

# Sample Stations Plot
ggplot() +
  geom_sf(data = USFE.riskregions) +
  geom_sf(data = st.df, aes(color = Subregion)) +
  scale_color_brewer(palette = "Set1") + # not color-blind safe
  ggtitle("Ceden Benthic Sample Stations")

```

## Water Quality Data

```{r, echo=TRUE, results='hide'}

# Bring in WQ data
ceden.wq <- read_excel("data/ceden_wq.xlsx")

#Remove records without lat/lon
ceden.wq <- ceden.wq[!is.na(ceden.wq$TargetLatitude),]

# Excel Data to sf Object: needs x, y, CRS
ceden.wq <- ceden.wq %>%
  st_as_sf(coords=c("TargetLongitude", "TargetLatitude"), crs = crs.WGS84)

# Spatial Join to add risk regions
ceden.wq.sf <- st_join(ceden.wq, USFE.riskregions["Subregion"])

# Remove records outside of Risk Regions
ceden.wq.sf <- ceden.wq.sf %>%
  filter(!is.na(Subregion))

```


```{r, echo=TRUE}

# Basic plot
ggplot() +
  geom_sf(data = USFE.riskregions) +
  geom_sf(data = ceden.wq.sf, aes(color = Subregion)) +
  scale_color_brewer(palette = "Set1") + # not color-blind safe
  ggtitle("Ceden WQ Data")

```


## Organize Water Quality Data

I created a dataframe that grouped by station name and sample date where each record is a water quality sampling event. This way the water quality sampling events can be matched up the the MI sampling events by location and timing.

This dataframe summarized water quality parameters by number of samples during that day, mean, standard deviation, minimum value, and maximum value. Though most of the parameters only had one measurement for the day.

The parameters included were:
* Alkalinity as CaCO3, Total
* Ammonia as N, Total
* Chlorine, Free, Total
* Chlorine, Total Residue, Total
* Oxygen, Dissolved, Total
* pH
* Salinity, Total
* Secchi Depth
* Specific Conductivity, Total
* Temperature
* Turbidity, Total
* Velocity

```{r, echo=TRUE, results='hide'}

wq.stations <- ceden.wq.sf %>%
  group_by(StationName, SampleDate) %>%
  summarise(Project = first(Project),
            n = n(),
            Subregion = first(Subregion),
            
            n_Alk = sum(Analyte == "Alkalinity as CaCO3, Total", na.rm = TRUE),
            mean_Alk = mean(Result[Analyte == "Alkalinity as CaCO3, Total"], na.rm = TRUE),
            sd_Alk = sd(Result[Analyte == "Alkalinity as CaCO3, Total"], na.rm = TRUE),
            min_Alk = min(Result[Analyte == "Alkalinity as CaCO3, Total"], na.rm = TRUE),
            max_Alk = max(Result[Analyte == "Alkalinity as CaCO3, Total"], na.rm = TRUE),

            n_N = sum(Analyte == "Ammonia as N, Total", na.rm = TRUE),
            mean_N = mean(Result[Analyte == "Ammonia as N, Total"], na.rm = TRUE),
            sd_N = sd(Result[Analyte == "Ammonia as N, Total"], na.rm = TRUE),
            min_N = min(Result[Analyte == "Ammonia as N, Total"], na.rm = TRUE),
            max_N = max(Result[Analyte == "Ammonia as N, Total"], na.rm = TRUE),
            
            n_Chl_F = sum(Analyte == "Chlorine, Free, Total", na.rm = TRUE),
            mean_Chl_F = mean(Result[Analyte == "Chlorine, Free, Total"], na.rm = TRUE),
            sd_Chl_F = sd(Result[Analyte == "Chlorine, Free, Total"], na.rm = TRUE),
            min_Chl_F = min(Result[Analyte == "Chlorine, Free, Total"], na.rm = TRUE),
            max_Chl_F = max(Result[Analyte == "Chlorine, Free, Total"], na.rm = TRUE),
            
            n_Chl_TR = sum(Analyte == "Chlorine, Total Residual, Total", na.rm = TRUE),
            mean_Chl_TR = mean(Result[Analyte == "Chlorine, Total Residual, Total"], na.rm = TRUE),
            sd_Chl_TR = sd(Result[Analyte == "Chlorine, Total Residual, Total"], na.rm = TRUE),
            min_Chl_TR = min(Result[Analyte == "Chlorine, Total Residual, Total"], na.rm = TRUE),
            max_Chl_TR = max(Result[Analyte == "Chlorine, Total Residual, Total"], na.rm = TRUE),
            
            n_DO = sum(Analyte == "Oxygen, Dissolved, Total", na.rm = TRUE),
            mean_DO = mean(Result[Analyte == "Oxygen, Dissolved, Total"], na.rm = TRUE),
            sd_DO = sd(Result[Analyte == "Oxygen, Dissolved, Total"], na.rm = TRUE),
            min_DO = min(Result[Analyte == "Oxygen, Dissolved, Total"], na.rm = TRUE),
            max_DO = max(Result[Analyte == "Oxygen, Dissolved, Total"], na.rm = TRUE),
            
            n_pH = sum(Analyte == "pH", na.rm = TRUE),
            mean_pH = mean(Result[Analyte == "pH"], na.rm = TRUE),
            sd_pH = sd(Result[Analyte == "pH"], na.rm = TRUE),
            min_pH = min(Result[Analyte == "pH"], na.rm = TRUE),
            max_pH = max(Result[Analyte == "pH"], na.rm = TRUE),
            
            n_Sal = sum(Analyte == "Salinity, Total", na.rm = TRUE),
            mean_Sal = mean(Result[Analyte == "Salinity, Total"], na.rm = TRUE),
            sd_Sal = sd(Result[Analyte == "Salinity, Total"], na.rm = TRUE),
            min_Sal = min(Result[Analyte == "Salinity, Total"], na.rm = TRUE),
            max_Sal = max(Result[Analyte == "Salinity, Total"], na.rm = TRUE),
            
            n_Secc = sum(Analyte == "Secchi Depth", na.rm = TRUE),
            mean_Secc = mean(Result[Analyte == "Secchi Depth"], na.rm = TRUE),
            sd_Secc = sd(Result[Analyte == "Secchi Depth"], na.rm = TRUE),
            min_Secc = min(Result[Analyte == "Secchi Depth"], na.rm = TRUE),
            max_Secc = max(Result[Analyte == "Secchi Depth"], na.rm = TRUE),
            
            n_Cond = sum(Analyte == "SpecificConductivity, Total", na.rm = TRUE),
            mean_Cond = mean(Result[Analyte == "SpecificConductivity, Total"], na.rm = TRUE),
            sd_Cond = sd(Result[Analyte == "SpecificConductivity, Total"], na.rm = TRUE),
            min_Cond = min(Result[Analyte == "SpecificConductivity, Total"], na.rm = TRUE),
            max_Cond = max(Result[Analyte == "SpecificConductivity, Total"], na.rm = TRUE),
            
            n_Temp = sum(Analyte == "Temperature", na.rm = TRUE),
            mean_Temp = mean(Result[Analyte == "Temperature"], na.rm = TRUE),
            sd_Temp = sd(Result[Analyte == "Temperature"], na.rm = TRUE),
            min_Temp = min(Result[Analyte == "Temperature"], na.rm = TRUE),
            max_Temp = max(Result[Analyte == "Temperature"], na.rm = TRUE),
            
            n_Turb = sum(Analyte == "Turbidity, Total", na.rm = TRUE),
            mean_Turb = mean(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            sd_Turb = sd(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            min_Turb = min(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            max_Turb = max(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            
            n_Turb = sum(Analyte == "Turbidity, Total", na.rm = TRUE),
            mean_Turb = mean(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            sd_Turb = sd(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            min_Turb = min(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            max_Turb = max(Result[Analyte == "Turbidity, Total"], na.rm = TRUE),
            
            n_Vel = sum(Analyte == "Velocity", na.rm = TRUE),
            mean_Vel = mean(Result[Analyte == "Velocity"], na.rm = TRUE),
            sd_Vel = sd(Result[Analyte == "Velocity"], na.rm = TRUE),
            min_Vel = min(Result[Analyte == "Velocity"], na.rm = TRUE),
            max_Vel = max(Result[Analyte == "Velocity"], na.rm = TRUE),
            
            ) %>%
  rename(SampleDate.wq = SampleDate)

### Change infinities and NaN values to NA
wq.stations <- wq.stations %>% 
  mutate_if(is.numeric, list(~na_if(., Inf))) %>% 
  mutate_if(is.numeric, list(~na_if(., -Inf))) %>%
  mutate_if(is.numeric, list(~na_if(., "NaN"))) %>%
  mutate_if(is.numeric, list(~na_if(., NaN)))

```

```{r, echo=TRUE}

tibble(wq.stations)

```

## Combine Benthic MI data and Water Quality Data

### Transform Projection to UTM Zone 10n

This transform allows us to compare distances and create buffers. I created a 500 meter buffer around each water quality sampling station.

```{r, echo=TRUE, results='hide'}

####### Transform into projection to compare distance

# Transform into UTM Zone 10n EPSG:26910
wq.stations <- st_transform(wq.stations, 26910)
st.df.u10 <- st_transform(st.df, 26910)
rr.u10 <- st_transform(USFE.riskregions, 26910)
samp.df.u10 <- st_transform(samp.df, 26910)

### Create 500m buffer around WQ sampling locations
wq.stations.buffer <- st_buffer(wq.stations, 500) # buffer is 500 meters

### Remove any buffers outside of the risk regions
wq.stations.buffer <- wq.stations.buffer %>%
  filter(!is.na(Subregion))

```

### Plot WQ buffers and ceden benthic data

```{r, echo=TRUE}
ggplot() +
  geom_sf(data = USFE.riskregions) +
  geom_sf(data = wq.stations.buffer, aes(color = Subregion)) +
  scale_color_brewer(palette = "Set1") + # not color-blind safe
  geom_sf(data = st.df.u10) +
  ggtitle("Ceden WQ Buffers and MI Sampling Locations")

```

### Join Datasets

This join adds the water quality sampling data to the MI sampling data where the MI sampling location falls within the 500 meter buffer around the water quality sampling location. The second line of code selects the records only where the sampling dates are the same.

```{r, echo=TRUE, results='hide'}

### Join Benthic locations and WQ buffers
samp.wq.com <- st_join(samp.df.u10, wq.stations.buffer, left = TRUE)

### Select records that have sampling data from the same date for benthic and WQ
com.dates <- samp.wq.com %>%
  filter(SampleDate == SampleDate.wq)

```

```{r, echo=TRUE}
### Plot benthic sampling locations that has WQ available on the same date

ggplot() +
  geom_sf(data = USFE.riskregions) +
  geom_sf(data = com.dates, aes(color = Subregion.x)) +
  scale_color_brewer(palette = "Set1") + # not color-blind safe
  ggtitle("Ceden Benthic Data with WQ on same date")

tibble(com.dates)
```


## Analysis






